---
output: html_document
params:
   biomfile: 
      value: x
   samplefile:
      value: x
   treefile:
      value: x
   normalization:
      value: x
   outputRdata:
      value: x
   ranks : 
      value: x
---
<meta http-equiv="Content-Type" content="text/html; charset=UTF-8">
<style type="text/css">
.button {
    background-color: #2a7ecc; /* Blue */
    border          : none;
    color           : white;
    padding         : 5px 10px;
    text-align      : center;
    text-decoration : none;
    display         : inline-block;
    font-size       : 12px;
    margin          : 4px 2px;
    cursor          : pointer;
    border-radius   : 8px;
}
h4 { 
    display      : block;
    font-size    : 1em;
    margin-top   : 1.33em;
    margin-bottom: 1.33em;
    margin-left  : 0;
    margin-right : 0;
    font-weight  : bold;
    color        : #2a7ecc;
}
</style>

<script src="https://s3-us-west-2.amazonaws.com/s.cdpn.io/14082/FileSaver.js"></script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
```
<h3>FROGS Phyloseq: Import Data</h3>

 Phyloseq `r packageVersion("phyloseq")`
 
```{r include=FALSE, cache=FALSE}
library(phyloseq)
library(ggplot2)
#library(biom)
library(ape)

## import data
biomfile <- params$biomfile
data     <- import_biom(biomfile)
tax      <- tax_table(data)[[1]]

if ((gregexpr('k__', tax))[[1]][1]>0) { 
    data <- import_biom(biomfile, parseFunction = parse_taxonomy_greengenes)
} else {
## change taxonomic rank names
    new_rank <- as.list(strsplit(params$ranks, " ")[[1]])
    colnames(tax_table(data)) <- new_rank
}

sampledata <- try(read.csv(params$samplefile, sep = "", row.names = 1), TRUE)
if (inherits(sampledata, "try-error") || ncol(sampledata)==0) {
    sampledata <- read.csv(params$samplefile, sep = "\t", row.names = 1)
}

sample_data(data) <- sampledata

if (params$treefile!="None"){
    treefile <- read.tree(params$treefile)
    phy_tree(data) <- treefile
}

## levels order as input sample_data file
for ( variable in sample_variables(data)){
  variable.order = as.vector(unique(sampledata[,variable]))
  sample_data(data)[,variable] <- factor(get_variable(data, variable),levels=variable.order)
}

## normalization
if(params$normalization){ data <- rarefy_even_depth(data, rngseed = 1121983)}

## save data
save(data, file=params$outputRdata)
```

#{.tabset .tabset-fade .tabset-pills}
##Summary
```{r,cache=FALSE, echo=FALSE, comment=NA}
data
if(params$normalization){cat(paste('Sequence number in each sample after normalization: ', head(sample_sums(data))[[1]]))}
```

##Ranks Names
```{r,cache=FALSE, echo=FALSE, comment=NA}
if ((gregexpr('k__', tax))[[1]][1]>0) {
  cat("Warning : Taxonomic affiliations come from Greengenes database, user specified ranks names are ignored.")
}
cat(paste('Rank names : ',paste(rank_names(data),collapse=', ')))
```

##Sample metadata
```{r,cache=FALSE, echo=FALSE, comment=NA}
variables <- sample_variables(data)
cat(paste('Sample variables: ', paste(variables, collapse=', ' )))

for (var in variables){
  cat(paste(var,': ',paste(levels(factor(get_variable(data, varName = var))),collapse=', '), '\n\n'))
}
```

##Plot tree
```{r, fig.width=10, fig.height=8, cache=FALSE, echo=FALSE, comment=NA}

if(params$treefile!="None"){
   p <- plot_tree(data, color=rank_names(data)[2]) + ggtitle(paste("Phylogenetic tree colored by", rank_names(data)[2])) + theme(plot.title = element_text(hjust = 0.5))
   plot(p)
}

if(params$treefile=="None"){
   cat("There is no phylogenetic tree in the object you have provided.")
}
```
<button id="dwn-plot-tree" class="button">
Download image
</button>


##R code
<h4>Loading packages</h4>
```{r, eval=FALSE}
library(phyloseq)
library(ape)
```

<h4>Importing data</h4>
```{r, eval=FALSE}
biomfile      <- # path to biom file
treefile      <- # path to newick file if possible
samplefile    <- # path to sample tsv file
outputRdata   <- # path to output Rdata file
normalization <- # TRUE or FALSE
ranks         <- # ranks names, ex: ranks <- "Royaume Phylum Classe Ordre Famille Genre Espece"

data <- import_biom(biomfile)
tax  <- tax_table(data)[[1]]

if ((gregexpr('k__', tax))[[1]][1]>0) { 
    data <-import_biom(biomfile, parseFunction = parse_taxonomy_greengenes)
} else {
## change taxonomic rank names
    new_rank <- as.list(strsplit(ranks, " ")[[1]])
    colnames(tax_table(data)) <- new_rank
}
sampledata <- try(read.csv(samplefile, sep = "", row.names = 1), TRUE)
if (inherits(sampledata, "try-error") || ncol(sampledata)==0) {
    sampledata <- read.csv(samplefile, sep = "\t", row.names = 1)
}

sample_data(data) <- sampledata
treefile          <- read.tree(treefile) #if have treefile
phy_tree(data)    <- treefile

# levels order as input sample_data file
for ( variable in sample_variables(data)){
  variable.order = as.vector(unique(sampledata[,variable]))
  sample_data(data)[,variable] <- factor(get_variable(data, variable),levels=variable.order)
}

# normalization
if(normalization) {
    data <- rarefy_even_depth(data, rngseed = 1121983);
    cat(paste('Sequence number in each sample after normalization: ', head(sample_sums(data))[[1]], "\n"))
}
```

<h4>Saving data</h4>
```{r, eval=FALSE}
save(data, file=outputRdata)
```

<h4>Data view</h4>
```{r, eval=FALSE}
data
## if have treefile
p <- plot_tree(data, color=rank_names(data)[2]) + ggtitle(paste("Phylogenetic tree colored by", rank_names(data)[2])) + theme(plot.title = element_text(hjust = 0.5))
plot(p)
```
<script>
header('Content-Type: text/html; charset=utf-8');
</script>

<div id="code-r-file" style="display:none;" charset=UTF-8>
library(phyloseq)
&#10;library(ape)
&#10;
&#10;## import data
&#10;biomfile <- # path to biom file
&#10;treefile <- # path to newick file if possible
&#10;samplefile <- # path to sample tsv file
&#10;outputRdata <- # path to output Rdata file
&#10;normalization <- # TRUE or FALSE
&#10;ranks <- # ranks names, ex: ranks <- \"Royaume Phylum Classe Ordre Famille Genre Espece\"
&#10;
&#10;data <- import_biom(biomfile)
&#10;tax <- tax_table(data)[[1]]
&#10;&#10;if ((gregexpr(\'k__\', tax))[[1]][1]>0) { 
&#10;    data <-import_biom(biomfile, parseFunction = parse_taxonomy_greengenes)
&#10;} else {
&#10;## change taxonomic rank names
&#10;    new_rank <- as.list(strsplit(ranks, \"&#32;\")[[1]])
&#10;    colnames(tax_table(data)) <- new_rank
&#10;}
&#10;sampledata <- try(read.csv(samplefile, sep = \"\", row.names = 1), TRUE)
&#10;if (inherits(sampledata, \"try-error\") || ncol(sampledata)==0) {
&#10;    sampledata <- read.csv(samplefile, sep = \"&#92;t\", row.names = 1)
&#10;}
&#10;sample_data(data) <- sampledata
&#10;
&#10;#if you have treefile
&#10;treefile <- read.tree(treefile)
&#10;phy_tree(data) <- treefile
&#10;
&#10;# levels order as input sample_data file
&#10;for ( variable in sample_variables(data)){
&#10;  variable.order = as.vector(unique(sampledata[,variable]))
&#10;  sample_data(data)[,variable] <- factor(get_variable(data, variable),levels=variable.order)
&#10;}
&#10;
&#10;## normalization
&#10;if(normalization) { 
&#10;data <- rarefy_even_depth(data, rngseed = 1121983);
&#10;cat(paste(\'Sequence number in each sample after normalization: \', head(sample_sums(data))[[1]], \"&#92;n\"))
&#10;}
&#10;
&#10;## save data
&#10;save(data, file=outputRdata)
&#10;
&#10;## data view
&#10;data
&#10;#if have treefile
&#10;p <- plot_tree(data, color=rank_names(data)[2]) + ggtitle(paste(\"Phylogenetic tree colored by\", rank_names(data)[2])) + theme(plot.title = element_text(hjust = 0.5))
&#10;plot(p)
</div>

<script>
  function download(){
    var text = document.getElementById("code-r-file").textContent;
    var blob = new Blob([text],{type:"text/plain;charset=utf-8"});
    saveAs(blob,"import_data_code.R");
   }
</script>
<input type="button" value="Download R code" class="button" onclick="download();"/>


<script>
function download1(filename, mysrc) {
var element = document.createElement('a');
    element.setAttribute('href', mysrc);
    element.setAttribute('download', filename);
    element.style.display = 'none';
    document.body.appendChild(element);
    element.click();
    document.body.removeChild(element);
}
document.getElementById("dwn-plot-tree").addEventListener("click", function(){
var div = document.getElementById('plot-tree');
var myimg = div.getElementsByTagName('img')[0];
var mysrc = myimg.src;
var filename = "tree_plot.png";
download1(filename, mysrc);
}, false);
</script>
