---
output: html_document
params:
   data:
      value: x
   varExp:
      value: x
   method:
      value: x
---
<meta http-equiv="Content-Type" content="text/html; charset=UTF-8">
<style type="text/css">
code.r{ /* Code block */
  font-size: 11px;
}
pre{ /* Code block */
  font-size: 11px
}
.button {
    background-color: #2a7ecc; /* Blue */
    border : none;
    color  : white;
    padding: 5px 10px;
    text-align: center;
    text-decoration: none;
    display  : inline-block;
    font-size: 12px;
    margin   : 4px 2px;
    cursor   : pointer;
    border-radius: 8px;
}
h4 { 
    display   : block;
    font-size : 1em;
    margin-top: 1.33em;
    margin-bottom: 1.33em;
    margin-left : 0;
    margin-right: 0;
    font-weight : bold;
    color : #2a7ecc;
}
</style>

<script src="https://s3-us-west-2.amazonaws.com/s.cdpn.io/14082/FileSaver.js"></script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<h3>FROGS Phyloseq: Multivariate Analysis of Variance</h3>

 Phyloseq `r packageVersion("phyloseq")`

```{r include=FALSE, echo=FALSE}
library(phyloseq)
library(vegan)
load(params$data)

##get dist object from distance matrix file
setwd(dirname(params$method))
A        <- read.table(file=basename(params$method), row.names=1)
dist.a   <- as.dist(A)
metadata <- as(sample_data(data), "data.frame") ## convert sample_data to data.frame
```

#{.tabset .tabset-fade .tabset-pills}
##Multivariate ANOVA performed with adonis
```{r, echo=FALSE, cach=FALSE, comment=NA}
adonis <- paste('adonis(dist.a ~ ', params$varExp, ', data = metadata, perm = 9999)')
eval(parse(text = adonis))
```

##Canonical Analysis of Principal Coordinates through capscale
```{r, echo=FALSE, cach=FALSE, comment=NA, warning=FALSE}
capcap <- paste('cap <- capscale(dist.a ~ ', params$varExp, ', data = metadata)')
eval(parse(text = capcap))
cap
anova(cap, permutations = 999)
```

##R code
<h4>Loading packages & data</h4>
```{r, eval=FALSE}
library(vegan)
## if in new session
# library(phyloseq)
#
## load phyloseq object
# data <- path to Rdata file
# load(data)
# extract sample metadata in a dataframe
metadata <- as(sample_data(data), "data.frame")

## load beta distance matrix
beta_file <- # path to Beta diversity tsv file
A <- read.table(file=beta_file, row.names=1)
beta_dist <- as.dist(A)
```
<h4>With adonis</h4>
```{r, comment=NA, cache=FALSE, echo=FALSE}
cat(paste(sep="","## change experiment variable (currently ", params$varExp, " )if you wish\n","## Multivariate ANOVA performed with adonis",'\nadonis(beta_dist ~ ', params$varExp, ', data = metadata, perm = 9999)'))
```

<h4>Through capscale</h4>
```{r, comment=NA, cache=FALSE, echo=FALSE}
cat(paste(sep="","## change experiment variable (currently ", params$varExp, " ) if you wish\n", "## Canonical Analysis of Principal Coordinates through capscale",'\ncap <- capscale(beta_dist ~ ', params$varExp, ', data = metadata)'), "\nanova(cap, permutations = 999)")
```

<div id="code-r-file" style="display:none;">
library(vegan)
&#10;#if in new session
&#10;#library(phyloseq)
&#10;#data <- path to Rdata file
&#10;#load(data)
&#10;metadata <- as(sample_data(data), \"data.frame\")
&#10;
&#10;varExp <- #Experiment variable
&#10;
&#10;beta_file <- # path to Beta diversity tsv file
&#10;A <- read.table(file=beta_file, row.names=1) 
&#10;beta_dist <- as.dist(A) 
&#10;
&#10;f <- paste(\"beta_dist ~ \" , varExp)
&#10;
&#10;## Multivariate ANOVA performed with adonis
&#10;adonis(as.formula(f), data = metadata, perm = 9999)
&#10;
&#10;## Canonical Analysis of Principal Coordinates through capscale
&#10;cap <- capscale(as.formula(f), data = metadata)
&#10;cap
&#10;anova(cap, permutations = 999)
</div>

<script>
  function download(){
    var text = document.getElementById("code-r-file").textContent;
    var blob = new Blob([text],{type:"text/plain;charset=utf-8"});
    saveAs(blob,"manova_code.R");
   }
</script>
<input type="button" value="Download R code" class="button" onclick="download();"/>
