---
output: html_document
params:
   data:
      value: x
   measures:
      value: x
   varExp:
      value: x
   fileAlpha:
      value: x
   libdir:
      value: x
---

<meta http-equiv="Content-Type" content="text/html; charset=UTF-8">
<style type="text/css">
code.r{ /* Code block */
  font-size: 11px;
}
pre { /* Code block */
  font-size: 11px
}
.button {
    background-color: #2a7ecc; /* Blue */
    border          : none;
    color           : white;
    padding         : 5px 10px;
    text-align      : center;
    text-decoration : none;
    display         : inline-block;
    font-size       : 12px;
    margin          : 4px 2px;
    cursor          : pointer;
    border-radius   : 8px;
}
h4 { 
    display      : block;
    font-size    : 1em;
    margin-top   : 1.33em;
    margin-bottom: 1.33em;
    margin-left  : 0;
    margin-right : 0;
    font-weight  : bold;
    color        : #2a7ecc;
}
</style>

<script src=`r file.path(params$libdir, "FileSaver.js")`></script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<h3>FROGS Phyloseq: Alpha Diversity Visualization</h3>

 Phyloseq `r packageVersion("phyloseq")`
 
```{r include=FALSE, cache=FALSE}
library(phyloseq)
library(ggplot2)
load(params$data)
measures <- as.list(strsplit(params$measures, ",")[[1]])
```

```{r, echo=FALSE}
## Numeric values of alpha diversity indices
alpha.diversity <- estimate_richness(data, measures = measures)

## export diversity to text file
write.table(alpha.diversity, params$fileAlpha, sep="\t", quote = FALSE, col.names = NA)
```

#{.tabset .tabset-fade .tabset-pills}
##Richness plot
```{r, fig.width=10, fig.height=8, echo=FALSE, comment=NA, warning=FALSE}
p <- plot_richness(data, x = params$varExp, color = params$varExp, measures = measures) + ggtitle(paste("Alpha diversity distribution in function of", params$varExp))+ theme(plot.title = element_text(hjust = 0.5))
plot(p)
```
<br>
<button id="dwn-richness-plot" class="button">
Download image
</button>

##Richness plot with boxplot
```{r, fig.width=10, fig.height=8, echo=FALSE, comment=NA, warning=FALSE}
p <- p + geom_boxplot(alpha = 0.2) +
         geom_point()+ theme_grey() +
         theme(axis.text.x = element_text(angle=90, hjust=0.5)) +
         theme(plot.title = element_text(hjust = 0.5))
plot(p)
```
<br>
<button id="dwn-richness-plot-with-boxplot" class="button">
Download image
</button>

##Alpha Diversity Indice Anova Analysis
```{r, echo=FALSE, comment=NA}
anova_data <- cbind(sample_data(data), alpha.diversity)
anova_data$Depth <- sample_sums(data)

variables <- paste(sep=" + ", "Depth", params$varExp )
## Perform ANOVA on observed richness, which effects are significant
for (m in measures){
    f <- paste(m," ~ ", variables)
    cat(sep = "", "###############################################################\n#Perform ANOVA on ",m,", which effects are significant\nanova.",m," <-aov( ",f,", anova_data)\nsummary(anova.",m,")\n")
    anova_res <- aov( as.formula(f), anova_data)
    res <- summary(anova_res)
    print(res)
    cat("\n\n")
}
```

##Rarefaction curves
```{r include=FALSE, cache=FALSE}
##code of Mahendra Mariadassou
library(parallel)
library(vegan)
## Rarefaction curve, ggplot style
ggrare <- function(physeq, step = 10, label = NULL, color = NULL, plot = TRUE, parallel = FALSE, se = TRUE) {
    ## Args:
    ## - physeq: phyloseq class object, from which abundance data are extracted
    ## - step: Step size for sample size in rarefaction curves
    ## - label: Default `NULL`. Character string. The name of the variable
    ##          to map to text labels on the plot. Similar to color option
    ##          but for plotting text.
    ## - color: (Optional). Default ‘NULL’. Character string. The name of the
    ##          variable to map to colors in the plot. This can be a sample
    ##          variable (among the set returned by
    ##          ‘sample_variables(physeq)’ ) or taxonomic rank (among the set
    ##          returned by ‘rank_names(physeq)’).
    ##
    ##          Finally, The color scheme is chosen automatically by
    ##          ‘link{ggplot}’, but it can be modified afterward with an
    ##          additional layer using ‘scale_color_manual’.
    ## - color: Default `NULL`. Character string. The name of the variable
    ##          to map to text labels on the plot. Similar to color option
    ##          but for plotting text.
    ## - plot:  Logical, should the graphic be plotted.
    ## - parallel: should rarefaction be parallelized (using parallel framework)
    ## - se:    Default TRUE. Logical. Should standard errors be computed. 
    ## require vegan
    x <- as(otu_table(physeq), "matrix")
    if (taxa_are_rows(physeq)) { x <- t(x) }

    ## This script is adapted from vegan `rarecurve` function
    tot <- rowSums(x)
    S   <- rowSums(x > 0)
    nr  <- nrow(x)

    rarefun <- function(i) {
        cat(paste("rarefying sample", rownames(x)[i]), sep = "\n")
        n <- seq(1, tot[i], by = step)
        if (n[length(n)] != tot[i]) {
            n <- c(n, tot[i])
        }
        y <- rarefy(x[i, ,drop = FALSE], n, se = se)
        if (nrow(y) != 1) {
	    rownames(y) <- c(".S", ".se")
            return(data.frame(t(y), Size = n, Sample = rownames(x)[i]))
        } else {
            return(data.frame(.S = y[1, ], Size = n, Sample = rownames(x)[i]))
        }
    }
    if (parallel) {
        out <- mclapply(seq_len(nr), rarefun, mc.preschedule = FALSE)
    } else {
        out <- lapply(seq_len(nr), rarefun)
    }
    df <- do.call(rbind, out)
    
    ## Get sample data 
    if (!is.null(sample_data(physeq, FALSE))) {
        sdf <- as(sample_data(physeq), "data.frame")
        sdf$Sample <- rownames(sdf)
        data <- merge(df, sdf, by = "Sample")
        labels <- data.frame(x = tot, y = S, Sample = rownames(x))
        labels <- merge(labels, sdf, by = "Sample")
    }
    
    ## Add, any custom-supplied plot-mapped variables
    if( length(color) > 1 ){
        data$color <- color
        names(data)[names(data)=="color"] <- deparse(substitute(color))
        color <- deparse(substitute(color))
    }
    if( length(label) > 1 ){
        labels$label <- label
        names(labels)[names(labels)=="label"] <- deparse(substitute(label))
        label <- deparse(substitute(label))
    }
    
    p <- ggplot(data = data, aes_string(x = "Size", y = ".S", group = "Sample", color = color))
    p <- p + labs(x = "Sample Size", y = "OTU Richness")
    if (!is.null(label)) {
        p <- p + geom_text(data = labels, aes_string(x = "x", y = "y", label = label, color = color),
                           size = 4, hjust = 0)
    }
    p <- p + geom_line()
    if (se) { ## add standard error if available
        p <- p + geom_ribbon(aes_string(ymin = ".S - .se", ymax = ".S + .se", color = NULL, fill = color), alpha = 0.2)
    }
    if (plot) {
        plot(p)
    }
    invisible(p)
}

p <- ggrare(data, step = 500, color = params$varExp, plot = FALSE)
rare.level <- sample_sums(data)[[1]]
```

```{r, fig.width=10, fig.height=8, echo=FALSE, comment=NA, warning=FALSE}
po <- paste('p + facet_wrap(~',params$varExp,') + geom_vline(xintercept = rare.level, color = "gray60")')
eval(parse(text = po))
plot(p)
```
<br>
<button id="dwn-rarefaction-curves" class="button">
Download image
</button>

##R code
<h4>Loading packages & data</h4>
```{r, comment=NA, cache=FALSE, echo=FALSE}
measures <- gsub(",", '","', paste('c("', params$measures, '")', sep=""))
cat("#if in new session \n#data <- path to Rdata file \n#load(data)\n#library(phyloseq)\n#library(ggplot2)")
cat(paste("varExp <- ", params$varExp))
cat(paste("measures <- ", measures))
cat("fileAlpha <- \"alpha_diversity.tsv\"")
```

<h4>Save alpha distance table</h4>
```{r, comment=NA, cache=FALSE, echo=FALSE}
cat(paste("fileAlpha <- #path to output Alpha diversity file\nalpha.diversity <- estimate_richness(data, measures = measures )"), "\nwrite.table(alpha.diversity, fileAlpha, sep='\\t', quote = FALSE, col.names = NA)")
```

<h4>Richness plot</h4>
```{r, comment=NA, cache=FALSE, echo=FALSE}
cat(paste(sep="","p <- plot_richness(data, x = ", '"', params$varExp,'"', ", color = ", '"', params$varExp, '"', ", measures = measures )", '+ ggtitle("Alpha diversity distribution in function of ', params$varExp, '")'),'+ theme(plot.title = element_text(hjust = 0.5))', '\nplot(p)')
```

<h4>With box plot</h4>
```{r, comment=NA, cache=FALSE, echo=FALSE}
cat('p <- p + geom_boxplot(alpha = 0.2) + theme_bw() + geom_point() + theme(axis.text.x = element_blank(), plot.title = element_text(hjust = 0.5))\nplot(p)')
```

<h4>Anova analysis</h4>
```{r, comment=NA, eval=FALSE}
anova_data <- cbind(sample_data(data), alpha.diversity)
anova_data$Depth <- sample_sums(data)
variables=c()
for (var in sample_variables(data) ) {
  l = length(levels(as.factor(get_variable(data,var))))
  if(l > 1 && l < nsamples(data)){
    variables <- cbind(variables,var)
  }
}
variables <- paste(sep=" + ", "Depth", paste(collapse =" + ", variables ))
## Perform ANOVA on observed richness, which effects are significant
for (m in measures){
    f <- paste(m," ~ ", variables)
    cat(sep = "", "###############################################################\n
                   #Perform ANOVA on ",m,", which effects are significant\n
                   anova.",m," <-aov( ",f,", anova_data)\n
                   summary(anova.",m,")\n")
    anova_res <- aov( as.formula(f), anova_data)
    res <- summary(anova_res)
    print(res)
    cat("\n\n")
}
```

<h4>The rarefaction curve</h4>
```{r, eval=FALSE}
##code of Mahendra Mariadassou for the rarefaction curve
library(parallel)
library(vegan)

## Rarefaction curve, ggplot style
ggrare <- function(physeq, step = 10, label = NULL, color = NULL, plot = TRUE, parallel = FALSE, se = TRUE) {
    ## Args:
    ## - physeq: phyloseq class object, from which abundance data are extracted
    ## - step: Step size for sample size in rarefaction curves
    ## - label: Default `NULL`. Character string. The name of the variable
    ##          to map to text labels on the plot. Similar to color option
    ##          but for plotting text.
    ## - color: (Optional). Default ‘NULL’. Character string. The name of the
    ##          variable to map to colors in the plot. This can be a sample
    ##          variable (among the set returned by
    ##          ‘sample_variables(physeq)’ ) or taxonomic rank (among the set
    ##          returned by ‘rank_names(physeq)’).
    ##
    ##          Finally, The color scheme is chosen automatically by
    ##          ‘link{ggplot}’, but it can be modified afterward with an
    ##          additional layer using ‘scale_color_manual’.
    ## - color: Default `NULL`. Character string. The name of the variable
    ##          to map to text labels on the plot. Similar to color option
    ##          but for plotting text.
    ## - plot:  Logical, should the graphic be plotted.
    ## - parallel: should rarefaction be parallelized (using parallel framework)
    ## - se:    Default TRUE. Logical. Should standard errors be computed. 
    ## require vegan
    x <- as(otu_table(physeq), "matrix")
    if (taxa_are_rows(physeq)) { x <- t(x) }

    ## This script is adapted from vegan `rarecurve` function
    tot <- rowSums(x)
    S <- rowSums(x > 0)
    nr <- nrow(x)

    rarefun <- function(i) {
        cat(paste("rarefying sample", rownames(x)[i]), sep = "\n")
        n <- seq(1, tot[i], by = step)
        if (n[length(n)] != tot[i]) {
            n <- c(n, tot[i])
        }
        y <- rarefy(x[i, ,drop = FALSE], n, se = se)
        if (nrow(y) != 1) {
	    rownames(y) <- c(".S", ".se")
            return(data.frame(t(y), Size = n, Sample = rownames(x)[i]))
        } else {
            return(data.frame(.S = y[1, ], Size = n, Sample = rownames(x)[i]))
        }
    }
    if (parallel) {
        out <- mclapply(seq_len(nr), rarefun, mc.preschedule = FALSE)
    } else {
        out <- lapply(seq_len(nr), rarefun)
    }
    df <- do.call(rbind, out)
    
    ## Get sample data 
    if (!is.null(sample_data(physeq, FALSE))) {
        sdf <- as(sample_data(physeq), "data.frame")
        sdf$Sample <- rownames(sdf)
        data <- merge(df, sdf, by = "Sample")
        labels <- data.frame(x = tot, y = S, Sample = rownames(x))
        labels <- merge(labels, sdf, by = "Sample")
    }
    
    ## Add, any custom-supplied plot-mapped variables
    if( length(color) > 1 ){
        data$color <- color
        names(data)[names(data)=="color"] <- deparse(substitute(color))
        color <- deparse(substitute(color))
    }
    if( length(label) > 1 ){
        labels$label <- label
        names(labels)[names(labels)=="label"] <- deparse(substitute(label))
        label <- deparse(substitute(label))
    }
    
    p <- ggplot(data = data, aes_string(x = "Size", y = ".S", group = "Sample", color = color))
    p <- p + labs(x = "Sample Size", y = "OTU Richness")
    if (!is.null(label)) {
        p <- p + geom_text(data = labels, aes_string(x = "x", y = "y", label = label, color = color),
                           size = 4, hjust = 0)
    }
    p <- p + geom_line()
    if (se) { ## add standard error if available
        p <- p + geom_ribbon(aes_string(ymin = ".S - .se", ymax = ".S + .se", color = NULL, fill = color), alpha = 0.2)
    }
    if (plot) {
        plot(p)
    }
    invisible(p)
}

p <- ggrare(data, step = 500, color = varExp, plot = FALSE)
rare.level <- sample_sums(data)[[1]]
capcap <- paste('p <- p + facet_wrap(~', varExp, ') + geom_vline(xintercept = rare.level, color = "gray60")')
eval(parse(text = capcap))
plot(p)
```

<div id="code-r-file" style="display:none;">
&#10;#if in new session 
&#10;#data <- path to Rdata file 
&#10;#load(data)
&#10;#library(phyloseq)
&#10;#library(ggplot2)
&#10;varExp <- # Experiment variable
&#10;measures <- # one list, ex:  c( \"Observed\",\"Chao1\",\"Shannon\",\"InvSimpson\")
&#10;#Save alpha distance table
&#10;fileAlpha <- #path to output Alpha diversity file
&#10;alpha.diversity <- estimate_richness(data, measures = measures ) 
&#10;write.table(alpha.diversity, fileAlpha,sep=\"\\t\", quote = FALSE, col.names = NA)
&#10;
&#10;#Richness plot
&#10;p <- plot_richness(data, x = varExp , color = varExp , measures = measures)+ ggtitle(\"Alpha diversity distribution\") + 
&#10;theme(plot.title = element_text(hjust = 0.5)) 
&#10;plot(p)
&#10;
&#10;#With box plot
&#10;p <- p + geom_boxplot(alpha = 0.2) + theme_bw() + geom_point() + theme(axis.text.x = element_blank(), plot.title = element_text(hjust = 0.5))
&#10;plot(p)
&#10;
&#10;#Anova analysis
&#10;anova_data <- cbind(sample_data(data), alpha.diversity)
&#10;anova_data$Depth <- sample_sums(data)
&#10;variables=c()
&#10;for (var in sample_variables(data) ) {
&#10;&#9; l = length(levels(as.factor(get_variable(data,var))))
&#10;&#9; if(l > 1 && l < nsamples(data)){
&#10;&#9; variables <- cbind(variables,var)
&#10;&#9; }
&#10;}
&#10;variables <- paste(sep=" + \", \"Depth" , paste(collapse = " + " , variables ))
&#10;#Perform anova analyses
&#10;for (m in measures){
&#10;&#9; f <- paste(m," ~ " , variables)
&#10;&#9; cat(sep = \"" , \"###############################################################\\n
&#10;&#9;&#9;&#9;&#9;&#9;&#9;&#9; #Perform ANOVA on " ,m,\", which effects are significant\\n
&#10;&#9;&#9;&#9;&#9;&#9;&#9;&#9; anova.\",m,\" <-aov( \",f,\", anova_data)\\n
&#10;&#9;&#9;&#9;&#9;&#9;&#9;&#9; summary(anova.\",m,\")\\n\")
&#10;&#9; anova_res <- aov( as.formula(f), anova_data)
&#10;&#9; res <- summary(anova_res)
&#10;&#9; print(res)
&#10;&#9; cat(\"\\n\\n\")
&#10;}
&#10;
&#10;##code of Mahendra Mariadassou for the rarefaction curve
&#10;library(parallel)
&#10;library(vegan)
&#10;## Rarefaction curve, ggplot style
&#10;ggrare <- function(physeq, step = 10, label = NULL, color = NULL, plot = TRUE, parallel = FALSE, se = TRUE) {
&#10;## Args:
&#10;## - physeq: phyloseq class object, from which abundance data are extracted
&#10;## - step: Step size for sample size in rarefaction curves
&#10;## - label: Default `NULL`. Character string. The name of the variable
&#10;##          to map to text labels on the plot. Similar to color option
&#10;##          but for plotting text.
&#10;## - color: (Optional). Default ‘NULL’. Character string. The name of the
&#10;##          variable to map to colors in the plot. This can be a sample
&#10;##          variable (among the set returned by
&#10;##          ‘sample_variables(physeq)’ ) or taxonomic rank (among the set
&#10;##          returned by ‘rank_names(physeq)’).
&#10;##
&#10;##          Finally, The color scheme is chosen automatically by
&#10;##          ‘link{ggplot}’, but it can be modified afterward with an
&#10;##          additional layer using ‘scale_color_manual’.
&#10;## - color: Default `NULL`. Character string. The name of the variable
&#10;##          to map to text labels on the plot. Similar to color option
&#10;##          but for plotting text.
&#10;## - plot:  Logical, should the graphic be plotted.
&#10;## - parallel: should rarefaction be parallelized (using parallel framework)
&#10;## - se:    Default TRUE. Logical. Should standard errors be computed. 
&#10;## require vegan
&#10;x <- as(otu_table(physeq), \"matrix\")
&#10;if (taxa_are_rows(physeq)) { x <- t(x) }
&#10;
&#10;## This script is adapted from vegan `rarecurve` function
&#10;tot <- rowSums(x)
&#10;S <- rowSums(x > 0)
&#10;nr <- nrow(x)
&#10;
&#10;rarefun <- function(i) {
&#10;        cat(paste(\"rarefying sample\", rownames(x)[i]), sep = \"&#92;n\")
&#10;        n <- seq(1, tot[i], by = step)
&#10;        if (n[length(n)] != tot[i]) {
&#10;            n <- c(n, tot[i])
&#10;        }
&#10;        y <- rarefy(x[i, ,drop = FALSE], n, se = se)
&#10;        if (nrow(y) != 1) {
&#10;	    rownames(y) <- c(\".S\", \".se\")
&#10;            return(data.frame(t(y), Size = n, Sample = rownames(x)[i]))
&#10;        } else {
&#10;            return(data.frame(.S = y[1, ], Size = n, Sample = rownames(x)[i]))
&#10;        }
&#10;    }
&#10;    if (parallel) {
&#10;        out <- mclapply(seq_len(nr), rarefun, mc.preschedule = FALSE)
&#10;    } else {
&#10;        out <- lapply(seq_len(nr), rarefun)
&#10;    }
&#10;df <- do.call(rbind, out)
&#10;    
&#10;## Get sample data 
&#10;if (!is.null(sample_data(physeq, FALSE))) {
&#10;sdf <- as(sample_data(physeq), \"data.frame\")
&#10;sdf&#36;Sample <- rownames(sdf)
&#10;data <- merge(df, sdf, by = \"Sample\")
&#10;labels <- data.frame(x = tot, y = S, Sample = rownames(x))
&#10;labels <- merge(labels, sdf, by = \"Sample\")
&#10;}
&#10;    
&#10;## Add, any custom-supplied plot-mapped variables
&#10;    if( length(color) > 1 ){
&#10;        data&#36;color <- color
&#10;        names(data)[names(data)==\"color\"] <- deparse(substitute(color))
&#10;        color <- deparse(substitute(color))
&#10;    }
&#10;    if( length(label) > 1 ){
&#10;        labels&#36;label <- label
&#10;        names(labels)[names(labels)==\"label\"] <- deparse(substitute(label))
&#10;        label <- deparse(substitute(label))
&#10;    }
&#10;    
&#10;    p <- ggplot(data = data, aes_string(x = \"Size\", y = \".S\", group = \"Sample\", color = color))
&#10;    p <- p + labs(x = \"Sample Size\", y = \"OTU Richness\")
&#10;    if (!is.null(label)) {
&#10;        p <- p + geom_text(data = labels, aes_string(x = \"x\", y = \"y\", label = label, color = color),
&#10;                           size = 4, hjust = 0)
&#10;    }
&#10;    p <- p + geom_line()
&#10;    if (se) { ## add standard error if available
&#10;        p <- p + geom_ribbon(aes_string(ymin = \".S - .se\", ymax = \".S + .se\", color = NULL, fill = color), alpha = 0.2)
&#10;    }
&#10;    if (plot) {
&#10;        plot(p)
&#10;    }
&#10;    invisible(p)
&#10;}
&#10;
&#10;p <- ggrare(data, step = 500, color = varExp, plot = FALSE)
&#10;rare.level <- sample_sums(data)[[1]]
&#10;
&#10;capcap <- paste(\'p <- p + facet_wrap(~\', varExp, \') + geom_vline(xintercept = rare.level, color = \"gray60\")\')
&#10;eval(parse(text = capcap))
&#10;plot(p)
&#10;
</div>

<script>
  function download1(){
    var text = document.getElementById("code-r-file").textContent;
    var blob = new Blob([text],{type:"text/plain;charset=utf-8"});
    saveAs(blob,"alpha_diversity_code.R");
   }
</script>
<input type="button" value="Download R code" class="button" onclick="download1();"/>

<script>
function download(filename, mysrc) {
var element = document.createElement('a');
    element.setAttribute('href', mysrc);
    element.setAttribute('download', filename);
    element.style.display = 'none';
    document.body.appendChild(element);
    element.click();
    document.body.removeChild(element);
}
document.getElementById("dwn-richness-plot").addEventListener("click", function(){
var div = document.getElementById('richness-plot');
var myimg = div.getElementsByTagName('img')[0];
var mysrc = myimg.src;
var filename = "richness_plot.png";
download(filename, mysrc);
}, false);

document.getElementById("dwn-richness-plot-with-boxplot").addEventListener("click", function(){
var div = document.getElementById('richness-plot-with-boxplot');
var myimg = div.getElementsByTagName('img')[0];
var mysrc = myimg.src;
var filename = "richness_plot_with_boxplot.png";
download(filename, mysrc);
}, false);

document.getElementById("dwn-rarefaction-curves").addEventListener("click", function(){
var div = document.getElementById('rarefaction-curves');
var myimg = div.getElementsByTagName('img')[0];
var mysrc = myimg.src;
var filename = "rarefaction_curves.png";
download(filename, mysrc);
}, false);
</script>
