---
output: html_document
params:
   data:
      value: x
   varExp:
      value: x
   method:
      value: x
---
<meta http-equiv="Content-Type" content="text/html; charset=UTF-8">

<style type="text/css">
code.r{ /* Code block */
  font-size: 11px;
}
pre { /* Code block */
  font-size: 11px
}
.button {
    background-color: #2a7ecc; /* Blue */
    border          : none;
    color           : white;
    padding         : 5px 10px;
    text-align      : center;
    text-decoration : none;
    display         : inline-block;
    font-size       : 12px;
    margin          : 4px 2px;
    cursor          : pointer;
    border-radius   : 8px;
}
h4 { 
    display      : block;
    font-size    : 1em;
    margin-top   : 1.33em;
    margin-bottom: 1.33em;
    margin-left  : 0;
    margin-right : 0;
    font-weight  : bold;
    color        : #2a7ecc;
}
</style>

<script src="https://s3-us-west-2.amazonaws.com/s.cdpn.io/14082/FileSaver.js"></script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<h3>FROGS Phyloseq: Sample clustering using different linkage method</h3>

 Phyloseq `r packageVersion("phyloseq")`

```{r include=FALSE, echo=FALSE}
library(phyloseq)
library(scales)
library(ape)
load(params$data)

##get dist object from distance matrix file
#setwd(dirname(params$method))
A <- read.table(file=params$method, row.names=1)
dist <- as.dist(A)

## Env types
var <- get_variable(data, params$varExp)

palette <- hue_pal()(length(levels(var))) # Automatic color palette: one color per different sample type

tipColor <- col_factor(palette, levels = levels(var))(var) # Map sample type to color
```

#{.tabset .tabset-fade .tabset-pills}

##Ward D2
```{r, fig.width=8, fig.height=6, echo=FALSE}
clust.wardD2 <- as.phylo(hclust(dist, method = 'ward.D2'))
par(mfcol = c(1,1))
plot(clust.wardD2, tip.color = tipColor, direction = "downwards", main = "Sample Clustering with Ward.D2 linkage")
legend("bottom", title = params$varExp, legend = levels(var) , xpd=TRUE, fill = palette, border = palette,cex=0.8, bty="n", ncol=4,  inset = c(0,-0.15))
```
<br>
<button id="dwn-ward-d2" class="button">
Download image
</button>

##Complete
```{r, fig.width=8, fig.height=6, echo=FALSE}
clust.complete <- as.phylo(hclust(dist, method = 'complete'))
par(mfcol = c(1,1))
plot(clust.complete, tip.color = tipColor, direction = "downwards", main = "Sample Clustering with Complete linkage")
legend("bottom", title = params$varExp, legend = levels(var) , xpd=TRUE, fill = palette, border = palette,cex=0.8, bty="n", ncol=4,  inset = c(0,-0.15))
```
<br>
<button id="dwn-complete" class="button">
Download image
</button>

##Single
```{r, fig.width=8, fig.height=6, echo=FALSE}
clust.single <- as.phylo(hclust(dist, method = 'single'))
par(mfcol = c(1,1))
plot(clust.single, tip.color = tipColor, direction = "downwards", main = "Sample Clustering with Single linkage")
legend("bottom", title = params$varExp, legend = levels(var) , xpd=TRUE, fill = palette, border = palette,cex=0.8, bty="n", ncol=4,  inset = c(0,-0.15))
```
<br>
<button id="dwn-single" class="button">
Download image
</button>

##R code

<h4>Loading packages & data</h4>
```{r, eval=FALSE}
library(scales)
library(ape)
#if in new session 
#library(phyloseq)
#data <- path to Rdata file 
#load(data)

##get dist object from distance matrix file
method <- # path to Beta diversity tsv file 
A      <- read.table(file=method, row.names=1)
dist   <- as.dist(A)

## Env types
varExp <- #Experiment variable to split plot
var    <- get_variable(data,varExp)
```

<h4>Make color palettes</h4>
```{r, eval=FALSE}
# automatic color palette: one color per different sample type
palette  <- hue_pal()(length(levels(var)))
# Map sample type to color
tipColor <- col_factor(palette, levels = levels(var))(var)
```

<h4>Sample clustering using different linkage method</h4>
```{r, eval=FALSE}
clust.wardD2 <- as.phylo(hclust(dist, method = 'ward.D2'))
par(mfcol = c(1,1))
plot(clust.wardD2, tip.color = tipColor, direction = "downwards", main = "Sample Clustering with Ward.D2 linkage")
legend("bottom", legend = levels(var) , xpd=TRUE, fill = palette, border = palette,cex=0.8, bty="n", ncol=4,  inset = c(0,-0.05))

clust.complete <- as.phylo(hclust(dist, method = 'complete'))
par(mfcol = c(1,1))
plot(clust.complete, tip.color = tipColor, direction = "downwards", main = "Sample Clustering with Complete linkage")
legend("bottom", legend = levels(var) , xpd=TRUE, fill = palette, border = palette,cex=0.8, bty="n", ncol=4,  inset = c(0,-0.05))

clust.single <- as.phylo(hclust(dist, method = 'single'))
par(mfcol = c(1,1))
plot(clust.single, tip.color = tipColor, direction = "downwards", main = "Sample Clustering with Single linkage")
legend("bottom", legend = levels(var) , xpd=TRUE, fill = palette, border = palette,cex=0.8, bty="n", ncol=4,  inset = c(0,-0.05))
```

<div id="code-r-file" style="display:none;" charset=UTF-8>
library(scales)
&#10;library(ape)
&#10;#if in new session 
&#10;#library(phyloseq)
&#10;#data <- path to Rdata file 
&#10;#load(data)
&#10;
&#10;##get dist object from distance matrix file
&#10;method <- # path to Beta diversity tsv file 
&#10;A <- read.table(file=method, row.names=1)
&#10;dist <- as.dist(A)
&#10;
&#10;## Env types
&#10;varExp <- #Experiment variable to split plot
&#10;var <- get_variable(data,varExp)
&#10;
&#10;## If samplefile have the Color column:
&#10;## Try this palette on clustering
&#10;## Map sample type to color
&#10;
&#10;# automatic color palette: one color per different sample type
&#10;palette <- hue_pal()(length(levels(var)))
&#10;# Map sample type to color
&#10;tipColor <- col_factor(palette, levels = levels(var))(var)
&#10;
&#10;clust.wardD2 <- as.phylo(hclust(dist, method = \'ward.D2\'))
&#10;par(mfcol = c(1,1))
&#10;plot(clust.wardD2, tip.color = tipColor, direction = \"downwards\", main = \"Sample Clustering with Ward.D2 linkage\")
&#10;
&#10;clust.complete <- as.phylo(hclust(dist, method = \'complete\'))
&#10;par(mfcol = c(1,1))
&#10;plot(clust.complete, tip.color = tipColor, direction = \"downwards\", main = \"Sample Clustering with Complete linkage\")
&#10;
&#10;clust.single <- as.phylo(hclust(dist, method = \'single\'))
&#10;par(mfcol = c(1,1))
&#10;plot(clust.single, tip.color = tipColor, direction = \"downwards\", main = \"Sample Clustering with Single linkage\")
&#10;
</div>

<script>
  function download1(){
    var text = document.getElementById("code-r-file").textContent;
    var blob = new Blob([text],{type:"text/plain;charset=utf-8"});
    saveAs(blob,"sample_clustering_code.R");
   }
</script>
<input type="button" value="Download R code" class="button" onclick="download1();"/>

<script>
function download(filename, mysrc) {
var element = document.createElement('a');
    element.setAttribute('href', mysrc);
    element.setAttribute('download', filename);
    element.style.display = 'none';
    document.body.appendChild(element);
    element.click();
    document.body.removeChild(element);
}
document.getElementById("dwn-ward-d2").addEventListener("click", function(){
    var div = document.getElementById('ward-d2');
    var myimg = div.getElementsByTagName('img')[0];
    var mysrc = myimg.src;
    var filename = "sample_clustering_wardD2.png";
    download(filename, mysrc);
}, false);

document.getElementById("dwn-complete").addEventListener("click", function(){
    var div = document.getElementById('complete');
    var myimg = div.getElementsByTagName('img')[0];
    var mysrc = myimg.src;
    var filename = "sample_clustering_complete.png";
    download(filename, mysrc);
}, false);

document.getElementById("dwn-single").addEventListener("click", function(){
    var div = document.getElementById('single');
    var myimg = div.getElementsByTagName('img')[0];
    var mysrc = myimg.src;
    var filename = "sample_clustering_single.png";
    download(filename, mysrc);
}, false);
</script>
