---
copyright: "Copyright (C) 2025 INRAE"
license: "GNU General Public License"
output: 
  html_notebook:
    code_folding: hide
params:
   phyloseq:
      value: x
   varExp:
      value: x
   method:
      value: x
   distance:
      value: x
   libdir:
      value: x
   version:
      value: x
   tool:
      value: x
---
<meta http-equiv="Content-Type" content="text/html; charset=UTF-8" />
<meta name="author" content="Ta Thi Ngan - SIGENAE/GABI & Maria Bernard - SIGENAE/GABI" />
<meta name="version" content="5.1.0" />
<meta name="copyright" content="Copyright (C) 2025 INRAE" />


```{r, echo=FALSE, results="asis"}
CURRENT_DIR <- knitr::opts_knit$get("output.dir")
if (is.null(CURRENT_DIR)) {
  CURRENT_DIR <- getwd()
}
THEME_DIR <- normalizePath(
  file.path(dirname(CURRENT_DIR), "static"),
  winslash = "/",
  mustWork = FALSE
)

js_theme_file <- normalizePath(
  file.path(THEME_DIR, "js", "theme.js"),
  winslash = "/",
  mustWork = FALSE
)

lines <- readLines(js_theme_file, warn = FALSE)

sep <- which(grepl("^//##", trimws(lines)))

if (length(sep) < 2) {
  stop("Tag not found in theme.js: '//##'")
}

js_code <- lines[(sep[1] + 1):(sep[2] - 1)]

cat("<script>\n")
cat(js_code, sep = "\n")
cat("\n</script>")
```

<div style="display:flex; align-items:center; gap:15px; margin-bottom:20px;">
  <img id="logo" src="data:image/png;base64," alt="FROGS logo" style="height:180px;">
  <h1 style="margin:0;">Stat: Structure Visualisation<i><span class="text-accent">(`r params$tool` v`r params$version`)</span></i></h1>
</div>

<script>

$(function () {
  update_logo(CURRENT_THEME);
});


</script>


<style type="text/css">
:root {
			--frogsColor: #03A5A8; 
			--frogsColorHover: hsl(from var(--frogsColor) h calc(s + 4) calc(l - 3) / 1);
			--frogsPreColor: hsl(from var(--frogsColor) h s l / 0.1);
		}
.text-accent {
  color: #595c5f;
}
.main-container{
	max-width: 75% !important;
}
h1{
	color: var(--frogsColor);
}
h2{
	color: var(--frogsColor);
}
a {
	color: var(--frogsColor);
}
a:hover{
	color: var(--frogsColorHover);
}
.nav-pills > li.active > a, .nav-pills > li.active > a:focus{
	color: #fff;
	background-color: var(--frogsColor);
	border-color: #dee2e6 #dee2e6 #fff;
}
.nav-pills > li.active > a:hover {
	background-color: var(--frogsColorHover);
}
.button {
    background-color: var(--frogsColor);
    border          : none;
    color           : white;
    padding         : 5px 10px;
    text-align      : center;
    text-decoration : none;
    display         : inline-block;
    font-size       : 12px;
    margin          : 4px 2px;
    cursor          : pointer;
    border-radius   : 8px;
}
h4 { 
    display      : block;
    font-size    : 1em;
    margin-top   : 1.33em;
    margin-bottom: 1.33em;
    margin-left  : 0;
    margin-right : 0;
    font-weight  : bold;
    color        : var(--frogsColor);
}
code.r{ /* Code block */
  font-size: 11px;
}
pre{
  font-size: 11px ;
  background-color: var(--frogsPreColor) !important;
}
.frogsversion{
	position: absolute;
	right: 5%;
	color: var(--frogsColor);
	font-style: italic;
	font-size: smaller;
	padding:8px;
}
.frogsversion > a{
	border: none !important;
	display: inline-block;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```


```{r import, message=FALSE, warning=FALSE}
## Import packages
library(phyloseq)
library(plotly)
source(file.path(params$libdir, "graphical_methods.R"))
## Alternative to source all extra function from a github repo
# source("https://raw.githubusercontent.com/mahendra-mariadassou/phyloseq-extended/master/load-extra-functions.R")

## Setting variables
  ## The Phyloseq object (format rdata)
  # phyloseq <- ""

  ## The beta diversity distance matrix file
  # distance <- ""

  ## The ordination method you want to use
  ## You can choose between "MDS" (for MDS/PCoA), "NMDS" or "DPCoA"
  # method <- ""

  ## The experiment variable that you want to analyse
  # varExp <- ""

## Create input and parameters dataframe
  # params <- data.frame( "phyloseq" = phylose, "distance" = distance, "method" = method, "varExp" = varExp)

## Load data
  ## the phyloseq object
load(params$phyloseq)

  ## the distance matrix file
A    <- read.table(file=params$distance, row.names=1)
dist <- as.dist(A)

## Add SampleID variable to physeq metadata
sample_data(data)$SampleID <- sample_names(data)

junk = capture.output(ord  <- ordinate(data, method = params$method, distance = dist),file=NULL)
var  <- get_variable(data, params$varExp)
```

# {.tabset .tabset-fade .tabset-pills}

```{=html}
<div class="row">
  <div style="float:right">
  <select id="themechoice"
          class="form-select form-select-sm"
          style="width: auto;"
          onchange="update_theme_Rmd(this.value)"
          aria-label="Switch theme">
    <option selected disabled value="">Switch theme</option>
    <option disabled value="DefaultTheme">Default</option>
    <option value="CoralTheme">Coral</option>
    <option value="GoldTheme">Gold</option>
    <option value="SteelTheme">Steel</option>
  </select>
  </div>
</div>
```

## Ordination plot

<h1 id="plot1">
```{r ord1, fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
## plot(p1)
p1 <- plot_ordination(data, ord, color = params$varExp) + 
                      geom_text(aes(label = SampleID), alpha = 0) + ## add invisible labels
                      theme_bw() + ggtitle(paste(params$method,"")) +
                      theme(plot.title = element_text(hjust = 0.5)) 
ggplotly(p1, tooltip = c("colour", "label"))
```
</h1>

<h1 id="plot2" style="display: none;">
```{r ord2, fig.width=8, fig.height=6, warning=FALSE}
## plot(p2)
p2 <- plot_samples(data, ord, color = params$varExp) +
                      geom_text(aes(label = SampleID), alpha = 0) + ## add invisible labels  
                      theme_bw() +
                      ggtitle(paste(params$method,"")) +
                      theme(plot.title = element_text(hjust = 0.5))

# reorder samples as in p1
p2$data[params$varExp] <- factor(p2$data[, params$varExp], levels = levels(p1$data[,params$varExp]))

ggplotly(p2, tooltip = c("colour", "label"))
```
</h1>

<h1 id="plot22" style="display: none;">
```{r ord22, fig.width=8, fig.height=6, warning=FALSE}
## p2 is obtained with plot samples see previously
g <- paste('p22 <- p2 +  stat_ellipse(aes(group =', params$varExp,'))')
eval(parse(text = g))
ggplotly(p22, tooltip = c("colour", "label"))
```
</h1>

<button class="button" onclick="document.getElementById('plot1').style.display='' ; document.getElementById('plot2').style.display='none'; document.getElementById('plot22').style.display='none' ">Without variable names</button>
<button class="button" onclick="document.getElementById('plot2').style.display='' ; document.getElementById('plot1').style.display='none'; document.getElementById('plot22').style.display='none' ">With variable names</button>
<button class="button" onclick="document.getElementById('plot22').style.display='' ; document.getElementById('plot2').style.display='none'; document.getElementById('plot1').style.display='none'">With ellipses</button>

## Heatmap plot
```{r heatmap, fig.width=9, fig.height=7,message=FALSE, warning=FALSE}
p3 <- plot_heatmap(data, method = params$method, distance = dist) + scale_fill_gradient2(low = "#ffff00", 
                                                mid = "#ffa500",
                                               high = "#ff0000", 
                                           na.value = "white",
                                              trans = log_trans(4), 
                                           midpoint = log(100, base = 4))
e  <- paste('~', params$varExp) 
p3 <- p3 + facet_grid(e, scales = "free_x") +
           theme(plot.title = element_text(hjust = 0.5)) +
           ggtitle(paste("Heatmap plot with", params$varExp)) +
           ylab("ASV")
plot(p3)
```

## Reproducibility token

```{r session, echo=TRUE, eval=TRUE}
sessioninfo::session_info()
```