---
output:
   html_document
params:
   data:
      value: x
   varExp:
      value: x
   taxaRank1:
      value: x
   taxaSet1:
      value: x
   taxaRank2:
      value: x
   numberOfTaxa:
      value: x
---
<meta http-equiv="Content-Type" content="text/html; charset=UTF-8">

<style type="text/css">
code.r{ /* Code block */
  font-size: 11px;
}
pre{ /* Code block */
  font-size: 11px
}
.button {
    background-color: #2a7ecc; /* Blue */
    border          : none;
    color           : white;
    padding         : 5px 10px;
    text-align      : center;
    text-decoration : none;
    display         : inline-block;
    font-size       : 12px;
    margin          : 4px 2px;
    cursor          : pointer;
    border-radius   : 8px;
}
h4 { 
    display      : block;
    font-size    : 1em;
    margin-top   : 1.33em;
    margin-bottom: 1.33em;
    margin-left  : 0;
    margin-right : 0;
    font-weight  : bold;
    color        : #2a7ecc;
}
</style>

<script src="https://s3-us-west-2.amazonaws.com/s.cdpn.io/14082/FileSaver.js"></script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<h3>FROGS Phyloseq: Visualize Data Composition</h3>

 Phyloseq `r packageVersion("phyloseq")`

```{r, include=FALSE, cache=FALSE}
library(plotly)
source("https://raw.githubusercontent.com/mahendra-mariadassou/phyloseq-extended/master/load-extra-functions.R") 
load(params$data)
```

#{.tabset .tabset-fade .tabset-pills}
##Bar plot
```{r, fig.width=10, fig.height=8, echo=FALSE, comment=NA, cache=FALSE}
e <- paste('facet_wrap(~', params$varExp,', scales = "free_x", nrow = 1)')
p <- plot_bar(data,fill= params$taxaRank2) + eval(parse(text = e)) + ggtitle(paste("Bar plot colored by", params$taxaRank2))+         theme(plot.title = element_text(hjust = 0.5))
plot(p)
```
<br>
<button id="dwn-bar-plot" class="button">
Download image
</button>

##Composition plot
```{r, fig.width=10, fig.height=8, echo=FALSE, comment=NA, cache=FALSE, warning=FALSE}
taxaSet1 <- unlist(strsplit(params$taxaSet1, " "))
p        <- plot_composition(data, params$taxaRank1, taxaSet1, params$taxaRank2, params$numberOfTaxa, fill=params$taxaRank2) 
p        <- p + eval(parse(text = e)) + theme(plot.title = element_text(hjust = 0.5))
gp <- ggplotly(p)
gp$x$layout$annotations[[1]]$x <- -0.08
gp$x$layout$margin$b = 100
gp$x$layout$margin$t = 100
gp$x$layout$margin$l = 75
gp
```

##R code
<h4>Loading packages & data</h4>
```{r, eval=FALSE}
library(plotly)
source("https://raw.githubusercontent.com/mahendra-mariadassou/phyloseq-extended/master/load-extra-functions.R")
#if in new session 
#data <- path to Rdata file 
#load(data)
```

<h4>Bar plot</h4>
```{r, comment=NA, cache=FALSE, echo=FALSE}
taxaSet1 <- paste0("c(\"", paste(unlist(strsplit(params$taxaSet1, " ")), collapse = "\", \""), "\")")
cat(paste(sep = "","p <- plot_bar(data,fill=", '"', params$taxaRank2, '"', ") + facet_wrap(~", params$varExp, ', scales = "free_x", nrow = 1) + ggtitle("', paste("Bar plot colored by", params$taxaRank2, '")'), '+ theme(plot.title = element_text(hjust = 0.5))', '\nplot(p)'))
```

<h4>Composition plot</h4>
```{r, comment=NA, cache=FALSE, echo=FALSE}
cat(paste(sep = "",'taxaSet1 <- unlist(strsplit(', params$taxaSet1, ', " "))\n', "p <- plot_composition(data,", '"', params$taxaRank1, '" ,taxaSet1 , "', params$taxaRank2, '" ,', params$numberOfTaxa, ', fill="', params$taxaRank2, '"', ") + eval(", parse(text = e),")"),'+ theme(plot.title = element_text(hjust = 0.5))', '\nggplotly(p)')
```

<div id="code-r-file" style="display:none; charset=UTF-8">
library(plotly)
&#10;source(\"https://raw.githubusercontent.com/mahendra-mariadassou/phyloseq-extended/master/load-extra-functions.R \")
&#10;# if in new session 
&#10;# data <- path to Rdata file 
&#10;# load(data)
&#10;
&#10;varExp <- #Experiment variable
&#10;taxaRank1 <- #Taxonomic rank at which to filter your data
&#10;taxaRank2 <- #Taxonomic rank used for aggregation
&#10;taxaSet1 <- #Taxa (at previously specified taxonomic rank) to keep in the dataset
&#10;numberOfTaxa <- #Number of most abundant taxa to keep

&#10;# Bar plot
&#10;e <- paste(\'facet_wrap(~\', varExp, \', scales = \"free_x\", nrow = 1)\')
&#10;p <- plot_bar(data,fill= taxaRank2) + eval(parse(text = e)) + ggtitle(paste(\"Bar plot colored by\", taxaRank2))+ theme(plot.title = element_text(hjust = 0.5))
&#10;plot(p)
&#10;
&#10;# composition plot
&#10;taxaSet1 <- unlist(strsplit(taxaSet1, \"&#32;\"))
&#10;p <- plot_composition(data, taxaRank1, taxaSet1, taxaRank2, numberOfTaxa, fill=taxaRank2) 
&#10;p <- p + eval(parse(text = e)) + theme(plot.title = element_text(hjust = 0.5))
&#10;ggplotly(p)
</div>

<script>
  function download1(){
    var text = document.getElementById("code-r-file").textContent;
    var blob = new Blob([text],{type:"text/plain;charset=utf-8"});
    saveAs(blob,"composition_code.R");
   }
</script>
<input type="button" value="Download R code" class="button" onclick="download1();"/>

<script>
function download(filename, mysrc) {
var element = document.createElement('a');
    element.setAttribute('href', mysrc);
    element.setAttribute('download', filename);
    element.style.display = 'none';
    document.body.appendChild(element);
    element.click();
    document.body.removeChild(element);
}
document.getElementById("dwn-bar-plot").addEventListener("click", function(){
var div = document.getElementById('bar-plot');
var myimg = div.getElementsByTagName('img')[0];
var mysrc = myimg.src;
var filename = "bar_plot.png";
download(filename, mysrc);
}, false);
</script>
