---
output:
  html_notebook:
    code_folding : hide
params:
  data:
    value: x
  varExp:
    value: x
  taxaRank1:
    value: x
  taxaSet1:
    value: x
  taxaRank2:
    value: x
  numberOfTaxa:
    value: x
  libdir:
    value: x
---
<meta http-equiv="Content-Type" content="text/html; charset=UTF-8">

<style type="text/css">
code.r{ /* Code block */
  font-size: 11px;
}
pre{ /* Code block */
  font-size: 11px
}
.button {
    background-color: #2a7ecc; /* Blue */
    border          : none;
    color           : white;
    padding         : 5px 10px;
    text-align      : center;
    text-decoration : none;
    display         : inline-block;
    font-size       : 12px;
    margin          : 4px 2px;
    cursor          : pointer;
    border-radius   : 8px;
}
h4 { 
    display      : block;
    font-size    : 1em;
    margin-top   : 1.33em;
    margin-bottom: 1.33em;
    margin-left  : 0;
    margin-right : 0;
    font-weight  : bold;
    color        : #2a7ecc;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<h3>FROGS Phyloseq: Visualize Data Composition (version 3.2)</h3>

 Phyloseq `r packageVersion("phyloseq")`

```{r package_import, message=FALSE}
library(plotly)
library(phyloseq)
source(file.path(params$libdir, "graphical_methods.R")) 
load(params$data)
```

#{.tabset .tabset-fade .tabset-pills}
##Bar plot
```{r bar_plot, fig.width=10, fig.height=8}
e <- paste('facet_wrap(~', params$varExp,', scales = "free_x", nrow = 1)')
p <- plot_bar(data,fill= rank_names(data)[2]) + eval(parse(text = e)) + ggtitle(paste("Bar plot colored by", rank_names(data)[2]))+         theme(plot.title = element_text(hjust = 0.5))
plot(p)
```

##Composition plot
```{r composion_plot, fig.width=10, fig.height=8}
taxaSet1 <- unlist(strsplit(params$taxaSet1, " "))
p        <- plot_composition(data, params$taxaRank1, taxaSet1, params$taxaRank2, params$numberOfTaxa, fill=params$taxaRank2) 
p        <- p + eval(parse(text = e)) + theme(plot.title = element_text(hjust = 0.5))
gp <- ggplotly(p, tooltip = c("x", "y", "fill"))
gp$x$layout$annotations[[1]]$x <- -0.08
gp$x$layout$margin$l = 75
gp$x$layout$margin$b = 100
gp$x$layout$margin$t = 100
gp
```