---
output: html_document
params:
   data:
      value: x
   varExp:
      value: x
   methods:
      value: x
---
<meta http-equiv="Content-Type" content="text/html; charset=UTF-8">
<style type="text/css">
code.r{ /* Code block */
  font-size: 11px;
}
pre { /* Code block */
  font-size: 11px
}
.button {
    background-color: #2a7ecc; /* Blue */
    border          : none;
    color           : white;
    padding         : 5px 10px;
    text-align      : center;
    text-decoration : none;
    display         : inline-block;
    font-size       : 12px;
    margin          : 4px 2px;
    cursor          : pointer;
    border-radius   : 8px;
}
h4 { 
    display      : block;
    font-size    : 1em;
    margin-top   : 1.33em;
    margin-bottom: 1.33em;
    margin-left  : 0;
    margin-right : 0;
    font-weight  : bold;
    color        : #2a7ecc;
}
</style>

<script src="https://s3-us-west-2.amazonaws.com/s.cdpn.io/14082/FileSaver.js"></script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<h3>FROGS Phyloseq: Beta Diversity Visualization</h3>

 Phyloseq `r packageVersion("phyloseq")`

```{r include=FALSE, cache=FALSE}
library(phyloseq)
library(ggplot2)
library(gridExtra)
library(reshape2)
source("https://raw.githubusercontent.com/mahendra-mariadassou/phyloseq-extended/master/load-extra-functions.R")
load(params$data) 
```

```{r include=FALSE, cache=FALSE, comment=NA}
## Generic method for distance
methods <- as.list(strsplit(params$methods, ",")[[1]])

## Order samples according to grouping variable
sampleOrder <- levels(reorder(sample_names(data), as.numeric(get_variable(data, params$varExp)))) 
```

#{.tabset .tabset-fade .tabset-pills}

##Dist as heatmap plot
```{r, fig.width=8, fig.height=6, echo=FALSE, warning=FALSE}

for (method in methods){
   if (method=="unifrac"){
      dist.uf <- distance(data, method = "unifrac") ## Unifrac
      uf <- as.matrix(dist.uf)
      write.table(uf, "Unifrac.tsv", sep="\t", quote=FALSE, col.names=NA)
      p1 <- plot_dist_as_heatmap(dist.uf, order = sampleOrder, title = "Unifrac") + 
        theme(plot.title = element_text(hjust = 0.5))
      plot(p1)
   }else if (method=="wunifrac"){
      dist.wuf <- distance(data, method = "wunifrac") ## Weighted Unifrac
      wuf <- as.matrix(dist.wuf)
      write.table(wuf, "wUnifrac.tsv",sep="\t", quote=FALSE, col.names=NA)
      p2 <- plot_dist_as_heatmap(dist.wuf, order = sampleOrder, title = "wUnifrac") + 
        theme(plot.title = element_text(hjust = 0.5))
      plot(p2)
   }else if (method=="bray"){
      dist.bc <- distance(data, method = "bray") ## Bray-Curtis
      bc <- as.matrix(dist.bc)
      write.table(bc, "Bray_Curtis.tsv",sep="\t", quote=FALSE, col.names=NA)
      p3 <- plot_dist_as_heatmap(dist.bc, order = sampleOrder, title = "Bray-Curtis") + 
        theme(plot.title = element_text(hjust = 0.5))
      plot(p3)
   }else if (method=="cc"){
      dist.jac <- distance(data, method = "cc") ## Jaccard
      jac <- as.matrix(dist.jac)
      write.table(jac, "Jaccard_binary.tsv",sep="\t", quote=FALSE, col.names=NA)
      p4 <- plot_dist_as_heatmap(dist.jac, order = sampleOrder, title = "Jaccard") + 
        theme(plot.title = element_text(hjust = 0.5))
      plot(p4)
   }else {
      dist.a <- distance(data, method = method)
      a <- as.matrix(dist.a)
      write.table(a, paste(sep="", method, ".tsv"), sep="\t", quote=FALSE, col.names=NA)
      pa <- plot_dist_as_heatmap(dist.a, order = sampleOrder, title = method) + 
        theme(plot.title = element_text(hjust = 0.5))
      plot(pa)
    }
}
```

##R code
<h4>Loading packages & data</h4>
```{r, eval=FALSE}
library(phyloseq)
library(ggplot2)
library(gridExtra)
library(reshape2)
source("https://raw.githubusercontent.com/mahendra-mariadassou/phyloseq-extended/master/load-extra-functions.R")
#if in new session 
#data <- path to Rdata file 
#load(data) 
varExp <- #Experiment variable to split plot
method <- #Beta diversity method name  #To show all available distances: distanceMethodList
fileBeta <- "beta_diversity.tsv"  # output file path
```

<h4>Dist as heatmap plot</h4>
```{r, eval=FALSE}
## Order samples according to grouping variable
sampleOrder <- levels(reorder(sample_names(data), as.numeric(get_variable(data, varExp)))) 

dist.a <- distance(data, method = method)

pa <- plot_dist_as_heatmap(dist.a, order = sampleOrder, title = method) + 
  theme(plot.title = element_text(hjust = 0.5))
plot(pa)
```

<h4>Save distance matrix</h4>
```{r, eval=FALSE}
a <- as.matrix(dist.a)
write.table(a, fileBeta, sep="\t", quote=FALSE, col.names=NA)
```

<div id="code-r-file" style="display:none;">
library(phyloseq)
&#10;library(ggplot2)
&#10;library(gridExtra)
&#10;library(reshape2)
&#10;source(\"https://raw.githubusercontent.com/mahendra-mariadassou/phyloseq-extended/master/load-extra-functions.R \")
&#10;#if in new session 
&#10;#data <- path to Rdata file 
&#10;#load(data) 
&#10;varExp <- #Experiment variable to split plot
&#10;method <- #Beta diversity method name
&#10;fileBeta <- #path to output Beta diversity file
&#10;#To show all available distances: distanceMethodList
&#10;
&#10;## Order samples according to grouping variable
&#10;sampleOrder <- levels(reorder(sample_names(data), as.numeric(get_variable(data, varExp)))) 
&#10;
&#10;dist.a <- distance(data, method = method)
&#10;
&#10;pa <- plot_dist_as_heatmap(dist.a, order = sampleOrder, title = method) + 
&#10;  theme(plot.title = element_text(hjust = 0.5))
&#10;plot(pa)
&#10;
&#10;a <- as.matrix(dist.a)
&#10;write.table(a, fileBeta , sep=\"\\t\", quote=FALSE, col.names=NA)
</div>

<script>
  function download(){
    var text = document.getElementById("code-r-file").textContent;
    var blob = new Blob([text],{type:"text/plain;charset=utf-8"});
    saveAs(blob,"beta_diversity_code.R");
   }
</script>
<input type="button" value="Download R code" class="button" onclick="download();"/>
