---
copyright: "Copyright (C) 2025 INRAE"
license: "GNU General Public License"
output:
  html_notebook:
    code_folding : hide
params:
  phyloseq:
    value: x
  varExp:
    value: x
  distance:
    value: x
  libdir:
    value: x
  themedir:
    value: x
  version:
    value: x
  tool:
    value: x
---
<meta http-equiv="Content-Type" content="text/html; charset=UTF-8" />
<meta name="author" content="Ta Thi Ngan - SIGENAE/GABI & Maria Bernard - SIGENAE/GABI" />
<meta name="version" content="5.1.0" />
<meta name="copyright" content="Copyright (C) 2025 INRAE" />

```{r, echo=FALSE, results="asis"}
THEME_DIR <- normalizePath(
  file.path(params$themedir),
  winslash = "/",
  mustWork = FALSE
)

js_theme_file <- normalizePath(
  file.path(THEME_DIR, "js", "theme.js"),
  winslash = "/",
  mustWork = FALSE
)

lines <- readLines(js_theme_file, warn = FALSE)

sep <- which(grepl("^//##", trimws(lines)))

if (length(sep) < 2) {
  stop("Tag not found in theme.js: '//##'")
}

js_code <- lines[(sep[1] + 1):(sep[2] - 1)]

cat("<script>\n")
cat(js_code, sep = "\n")
cat("\n</script>")
```

<div style="display:flex; align-items:center; gap:15px; margin-bottom:20px;">
  <img id="logo" src="data:image/png;base64," alt="FROGS logo" style="height:180px;">
  <h1 style="margin:0;">Stat: Multivariate Analysis of Variance<i><span class="text-accent">(`r params$tool` v`r params$version`)</span></i></h1>
</div>

<script>

$(function () {
  update_logo(CURRENT_THEME);
});


</script>

<style type="text/css">
:root {
			--frogsColor: #03A5A8; 
			--frogsColorHover: hsl(from var(--frogsColor) h calc(s + 4) calc(l - 3) / 1);
			--frogsPreColor: hsl(from var(--frogsColor) h s l / 0.1);
		}
.text-accent {
  color: #595c5f;
}
.main-container{
	max-width: 75% !important;
}
h1{
	color: var(--frogsColor);
}
h2{
	color: var(--frogsColor);
}
a {
	color: var(--frogsColor);
}
a:hover{
	color: var(--frogsColorHover);
}
.nav-pills > li.active > a, .nav-pills > li.active > a:focus{
	color: #fff;
	background-color: var(--frogsColor);
	border-color: #dee2e6 #dee2e6 #fff;
}
.nav-pills > li.active > a:hover {
	background-color: var(--frogsColorHover);
}
.button {
    background-color: var(--frogsColor)
    border          : none;
    color           : white;
    padding         : 5px 10px;
    text-align      : center;
    text-decoration : none;
    display         : inline-block;
    font-size       : 12px;
    margin          : 4px 2px;
    cursor          : pointer;
    border-radius   : 8px;
}
h4 { 
    display      : block;
    font-size    : 1em;
    margin-top   : 1.33em;
    margin-bottom: 1.33em;
    margin-left  : 0;
    margin-right : 0;
    font-weight  : bold;
    color        : var(--frogsColor);
}
code.r{ /* Code block */
  font-size: 11px;
}
pre{
  font-size: 11px ;
  background-color: var(--frogsPreColor) !important;
}
.frogsversion{
	position: absolute;
	right: 5%;
	color: var(--frogsColor);
	font-style: italic;
	font-size: smaller;
	padding:8px;
}
.frogsversion > a{
	border: none !important;
	display: inline-block;
}
#themechoice{
  border-radius: 0.25em;
  background-color: white;
  border: 1px solid #dee2e6;
  font-weight: 400;
  line-height: 1.5;
  padding: .375rem 2.25rem .375rem .75rem;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, warning = FALSE)
```


# {.tabset .tabset-fade .tabset-pills}

```{=html}
<div class="row">
  <div style="float:right">
  <select id="themechoice"
          class="form-select form-select-sm"
          style="width: auto;"
          onchange="update_theme_Rmd(this.value)"
          aria-label="Switch theme">
    <option selected disabled value="">Switch theme</option>
    <option disabled value="DefaultTheme">Default</option>
    <option value="CoralTheme">Coral</option>
    <option value="GoldTheme">Gold</option>
    <option value="SteelTheme">Steel</option>
  </select>
  </div>
</div>
```

## Multivariate Analysis of Variance

```{r import, message=FALSE}
## Import packages
library(phyloseq)
library(vegan)

## Setting variables
	## The Phyloseq object (format rdata)
	# phyloseq <- ""

	## The beta diversity distance matrix file
	# distance <- ""

	## The experiment variable that you want to analyse
	# varExp <- ""

## Create input and parameters dataframe
  # params <- data.frame( "phyloseq" = phylose, "distance" = distance, "varExp" = varExp)

## Load data
  # the phyloseq object, nammed data in FROGSSTAT Phyloseq Import data
load(params$phyloseq)
  # Convert sample_data to data.frame
metadata <- as(sample_data(data), "data.frame") 

  # the distance matrix file
A        <- read.table(file=params$distance, row.names=1)
dist     <- as.dist(A)

## Multivariate ANOVA performed with adonis
adonis <- paste('adonis(dist ~ ', params$varExp, ', data = metadata, perm = 9999)')
adonis_res <- eval(parse(text = adonis))
print(adonis_res$call)
print(adonis_res$aov.tab)
```

## Reproducibility token

```{r session, echo=TRUE, eval=TRUE}
sessioninfo::session_info()
```