---
title: "FROGS Phyloseq: Multivariate Analysis of Variance (version `r params$version`)"
copyright: "Copyright (C) 2024 INRAE"
license: "GNU General Public License"
version: "5.0.0"
output:
  html_notebook:
    code_folding : hide
params:
   phyloseq:
      value: x
   varExp:
      value: x
   distance:
      value: x
   libdir:
      value: x
   version:
      value: x
---
<meta http-equiv="Content-Type" content="text/html; charset=UTF-8" />
<meta name="author" content="Ta Thi Ngan - SIGENAE/GABI & Maria Bernard - SIGENAE/GABI" />
<meta name="version" content="5.0.0" />
<meta name="copyright" content="Copyright (C) 2024 INRAE" />

<style type="text/css">
:root {
			--frogsColor: #388697; 
			--frogsColorHover: rgb(12,10,62,0.8);
			--frogsColorShadow: #6c5b64;
			--frogsColor2: #FA8883;
			--frogsColor3: #EFC7C2;
			--frogsColor4: #737f75;
			--frogsColorZoomHighCharts: rgb(191,211,193,0.25);
			--frogsButtonColor: rgb(12,10,62,0.6);
			--frogsButtonBorderColor: rgb(12,10,62,0.25);
		}
h1{
	color: var(--frogsColor);
}
h2{
	color: var(--frogsColor);
}
a {
	color: var(--frogsColor);
}
a:hover{
	color: var(--frogsColorHover);
}
.nav-pills > li.active > a, .nav-pills > li.active > a:focus{
	color: #fff;
	background-color: var(--frogsColor);
	border-color: #dee2e6 #dee2e6 #fff;
}
.nav-pills > li.active > a:hover {
	background-color: var(--frogsColorHover);
}
.button {
    background-color: var(--frogsButtonColor)
    border          : none;
    color           : white;
    padding         : 5px 10px;
    text-align      : center;
    text-decoration : none;
    display         : inline-block;
    font-size       : 12px;
    margin          : 4px 2px;
    cursor          : pointer;
    border-radius   : 8px;
}
h4 { 
    display      : block;
    font-size    : 1em;
    margin-top   : 1.33em;
    margin-bottom: 1.33em;
    margin-left  : 0;
    margin-right : 0;
    font-weight  : bold;
    color        : var(--frogsColor);
}
code.r{ /* Code block */
  font-size: 11px;
}
pre{
  font-size: 11px ;
  background-color: rgba(191, 211, 193, 0.25) !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, warning = FALSE)
```
Phyloseq `r packageVersion("phyloseq")`

```{r import, message=FALSE}
## Import packages
library(phyloseq)
library(vegan)

## Setting variables
	## The Phyloseq object (format rdata)
	# phyloseq <- ""

	## The beta diversity distance matrix file
	# distance <- ""

	## The experiment variable that you want to analyse
	# varExp <- ""

## Create input and parameters dataframe
  # params <- data.frame( "phyloseq" = phylose, "distance" = distance, "varExp" = varExp)

## Load data
  # the phyloseq object, nammed data in FROGSSTAT Phyloseq Import data
load(params$phyloseq)
  # Convert sample_data to data.frame
metadata <- as(sample_data(data), "data.frame") 

  # the distance matrix file
A        <- read.table(file=params$distance, row.names=1)
dist     <- as.dist(A)

## Multivariate ANOVA performed with adonis
adonis <- paste('adonis(dist ~ ', params$varExp, ', data = metadata, perm = 9999)')
adonis_res <- eval(parse(text = adonis))
print(adonis_res$call)
print(adonis_res$aov.tab)
```
