---
output: html_document
params:
   data:
      value: x
   varExp:
      value: x
   method:
      value: x
   distance:
      value: x
---
<meta http-equiv="Content-Type" content="text/html; charset=UTF-8">
<style type="text/css">
code.r{ /* Code block */
  font-size: 11px;
}
pre { /* Code block */
  font-size: 11px
}
.button {
    background-color: #2a7ecc; /* Blue */
    border : none;
    color  : white;
    padding: 5px 10px;
    text-align : center;
    text-decoration : none;
    display: inline-block;
    font-size: 12px;
    margin: 4px 2px;
    cursor: pointer;
    border-radius: 8px;
}
h4 { 
    display   : block;
    font-size : 1em;
    margin-top: 1.33em;
    margin-bottom: 1.33em;
    margin-left  : 0;
    margin-right : 0;
    font-weight  : bold;
    color: #2a7ecc;
}
</style>

<script src="https://s3-us-west-2.amazonaws.com/s.cdpn.io/14082/FileSaver.js"></script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<h3>FROGS Phyloseq: Structure Visualization</h3>

 Phyloseq `r packageVersion("phyloseq")`


```{r include=FALSE, echo=FALSE}
source("https://raw.githubusercontent.com/mahendra-mariadassou/phyloseq-extended/master/load-extra-functions.R")
load(params$data)

#get dist object from distance matrix file
A    <- read.table(file=params$distance, row.names=1)
dist <- as.dist(A)
ord  <- ordinate(data, method = params$method, distance = dist)
var  <- get_variable(data, params$varExp)

# automatic color palette: one color per different sample type
palette <- hue_pal()(length(levels(var)))
va      <- params$varExp
names(palette) <- levels(sample_data(data)$va)

p1 <- plot_ordination(data, ord, color = params$varExp) + 
                      scale_color_manual(values = palette) +
                      theme_bw() + ggtitle(paste(params$method,"")) +
                      theme(plot.title = element_text(hjust = 0.5)) 
p2 <- plot_samples(data, ord, color = params$varExp) +
                      scale_color_manual(values = palette) + theme_bw() +
                      ggtitle(paste(params$method,"")) +
                      theme(plot.title = element_text(hjust = 0.5))
g <- paste('p22 <- p2 +  stat_ellipse(aes(group =', params$varExp,'))')
eval(parse(text = g))

p3 <- plot_heatmap(data) + scale_fill_gradient2(low = "#ffff00", 
                                                mid = "#ffa500",
                                               high = "#ff0000", 
                                           na.value = "white",
                                              trans = log_trans(4), 
                                           midpoint = log(100, base = 4))
e  <- paste('~', params$varExp) 
p3 <- p3 + facet_grid(e, scales = "free_x") +
           theme(plot.title = element_text(hjust = 0.5)) +
           ggtitle(paste("Heatmap plot with", params$varExp))
```

#{.tabset .tabset-fade .tabset-pills}
##Ordination plot

<h1 id="plot1">
```{r, fig.width=8, fig.height=6, echo=FALSE, cache=FALSE}
plot(p1)
```
<br>
<button id="dwn-plot1" class="button">
Download image
</button>
</h1>

<h1 id="plot2" style="display: none;">
```{r, fig.width=8, fig.height=6, echo=FALSE, cache=FALSE}
plot(p2)
```
<br>
<button id="dwn-plot2" class="button">
Download image
</button>
</h1>

<h1 id="plot22" style="display: none;">
```{r, fig.width=8, fig.height=6, echo=FALSE, cache=FALSE, comment=NA, message=FALSE}
plot(p22)
```
<br>
<button id="dwn-plot22" class="button">
Download image
</button>
</h1>

<button class="button" onclick="document.getElementById('plot1').style.display='' ; document.getElementById('plot2').style.display='none'; document.getElementById('plot22').style.display='none' ">No sample name</button>

<button class="button" onclick="document.getElementById('plot2').style.display='' ; document.getElementById('plot1').style.display='none'; document.getElementById('plot22').style.display='none' ">With sample name</button>

<button class="button" onclick="document.getElementById('plot22').style.display='' ; document.getElementById('plot2').style.display='none'; document.getElementById('plot1').style.display='none'">With ellipse</button>

##Heatmap plot
```{r, fig.width=9, fig.height=7, echo=FALSE, cache=FALSE, warning=FALSE}
plot(p3)
```
<br>
<button id="dwn-heatmap-plot" class="button">
Download image
</button>

##R code
<h4>Loading packages & data</h4>
```{r, eval=FALSE}
library(phyloseq)
source("https://raw.githubusercontent.com/mahendra-mariadassou/phyloseq-extended/master/load-extra-functions.R")
#if you want to generate interactiv plot, load plotly and use ggplotly() instead of plot()
#library(plotly)

#if in new session
#data <- path to Rdata file containing phyloseq object
#load(data)
#distance <- path to tsv file containing beta diversity distance matrix to use
# A <- read.table(file=distance, row.names=1)
# dist <- as.dist(A)
```
<h4>Ordianation plots </h4>
```{r, comment=NA, cache=FALSE, echo=FALSE}
cat(paste(sep="","ord <- ordinate(data, method =", '"', params$method,'"', ", distance = dist)"))

cat(paste(sep="","p1 <- plot_ordination(data, ord, color = ", '"', params$varExp, '"', ") + theme_bw() + ggtitle(", '"', params$method, '")', '+ theme(plot.title = element_text(hjust = 0.5))','\nplot(p1)'))

cat(paste(sep="","p2 <- plot_samples(data, ord, color = ", '"', params$varExp, '"', ") + theme_bw()+ ggtitle(", '"', params$method, '")', '+ theme(plot.title = element_text(hjust = 0.5))','\nplot(p2)'))

cat(paste(sep="",'p22 <- p2 +  stat_ellipse(aes(group = ', params$varExp,'))', '\nplot(p22)'))
```

<h4>Heatmap plot</h4>
```{r, comment=NA, cache=FALSE, echo=FALSE}
cat('# change heatmap color scale with color name or hexadecimal code. see http://tools.medialab.sciences-po.fr/iwanthue/ \nlowCol <- "#ffff00" # yellow\nmidCol <-  "#ffa500" # orange\nhighCol <- "#ff0000" # red\nnaCol <- "white" # white!')

cat(paste(sep="",'p3 <- plot_heatmap(data) + scale_fill_gradient2(low = lowCol, mid = midCol, high = highCol, na.value = naCol, trans = log_trans(4), midpoint = log(100, base = 4))+ facet_grid(~', params$varExp, ', scales = "free_x") + theme(plot.title = element_text(hjust = 0.5)) + ggtitle("Heatmap plot")','\nplot(p3)'))
```


<div id="code-r-file" style="display:none;">
# ######### Loading packages & data & parameter setting #########
&#10;library(phyloseq)
&#10;source(\"https://raw.githubusercontent.com/mahendra-mariadassou/phyloseq-extended/master/load-extra-functions.R" )
&#10;#if you want to generate interactiv plot, load plotly and use ggplotly() instead of plot()
&#10;#library(plotly)
&#10;
&#10;#if in new session, uncomment and complete
&#10;#data <- # path to Rdata file
&#10;#load(data)
&#10;
&#10;#distance <- # path to Beta diversity tsv file
&#10;#A <- read.table(file=distance, row.names=1)
&#10;#dist <- as.dist(A)
&#10;
&#10;# to be complete
&#10;varExp <- #Experiment variable
&#10;method <- \"DCA\" # ordination method, ex \"MDS\", \"NMDS\", \"DCA\" (default), \"CCA\", \"RDA\", \"CAP\", \"DPCoA\" or \"PCoA\" 
&#10;
&#10;######### Ordiantion plots #########
&#10;ord <- ordinate(data, method =method, distance = dist)
&#10;
&#10;p1 <- plot_ordination(data, ord, color = varExp) + theme_bw() + ggtitle(method)+ theme(plot.title = element_text(hjust = 0.5))
&#10;plot(p1)
&#10;
&#10;p2 <- plot_samples(data, ord, color = varExp) + theme_bw()+ ggtitle(method)+ theme(plot.title = element_text(hjust = 0.5))
&#10;plot(p2)
&#10;
&#10;g <- paste(\'p22 <- p2 +  stat_ellipse(aes(group =\', varExp,\'))\')
&#10;eval(parse(text = g))
&#10;plot(p22)
&#10;
&#10;######### Heatmap Plot #########
&#10;# change heatmap color scale with color name or hexadecimal code. see http://tools.medialab.sciences-po.fr/iwanthue/
&#10;lowCol <- \"#ffff00\" # yellow
&#10;midCol <-  \"#ffa500\" # orange
&#10;highCol <- \"#ff0000\" # red
&#10;naCol <- \"white\" # white!
&#10;
&#10;p3 <- plot_heatmap(data) + scale_fill_gradient2(low = lowCol, mid = midCol, high = highCol, 
&#10;                              na.value = naCol, trans = log_trans(4), 
&#10;                              midpoint = log(100, base = 4))
&#10;e <- paste(\'~\', varExp) 
&#10;p3 <- p3 + facet_grid(e, scales = \"free_x\")+ theme(plot.title = element_text(hjust = 0.5))
&#10;plot(p3)
&#10;
</div>

<button id="dwn-btn" class="button">Download R code file</button>

<script>
function download1(filename, text) {
    var element = document.createElement('a');
    element.setAttribute('href', 'data:text/plain;charset=utf-8,' + encodeURIComponent(text));
    element.setAttribute('download', filename);

    element.style.display = 'none';
    document.body.appendChild(element);
    element.click();
    document.body.removeChild(element);
}

// Start file download.
document.getElementById("dwn-btn").addEventListener("click", function(){
    // Generate download of hello.txt file with some content
    var text = document.getElementById("code-r-file").innerText;
    var filename = "structure_visualization_code.R";
    download1(filename, text);
}, false);
</script>

<script>
function download(filename, mysrc) {
var element = document.createElement('a');
    element.setAttribute('href', mysrc);
    element.setAttribute('download', filename);
    element.style.display = 'none';
    document.body.appendChild(element);
    element.click();
    document.body.removeChild(element);
}
document.getElementById("dwn-plot1").addEventListener("click", function(){
    var div = document.getElementById('plot1');
    var myimg = div.getElementsByTagName('img')[0];
    var mysrc = myimg.src;
    var filename = "ordination_plot_no_sample_name.png";
    download(filename, mysrc);
}, false);

document.getElementById("dwn-plot2").addEventListener("click", function(){
    var div = document.getElementById('plot2');
    var myimg = div.getElementsByTagName('img')[0];
    var mysrc = myimg.src;
    var filename = "ordination_plot_with_sample_name.png";
    download(filename, mysrc);
}, false);

document.getElementById("dwn-plot22").addEventListener("click", function(){
    var div = document.getElementById('plot22');
    var myimg = div.getElementsByTagName('img')[0];
    var mysrc = myimg.src;
    var filename = "ordination_plot_with_sample_name_and_ellipse.png";
    download(filename, mysrc);
}, false);

document.getElementById("dwn-heatmap-plot").addEventListener("click", function(){
    var div = document.getElementById('heatmap-plot');
    var myimg = div.getElementsByTagName('img')[0];
    var mysrc = myimg.src;
    var filename = "heatmap_plot.png";
    download(filename, mysrc);
}, false);

</script>
